---
title: "EnorNOC-GAM"
author: "Yu-Chen Xue"
date: "2018年6月13日"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
opts_knit$set(root.dir = 'D:/WORKSPACE/RProjects/EnorNOC-GAM/all-data/csv')
```

## GAM methods are implemented in R in the package mgcv.

## For visualizations packages ggplot2, grid and animation will be used. One useful function from package car will be used too.

## Let’s scan all of the needed packages.


## 1. Dataset介紹：
名稱：EnerNOC GreenButton Data\
來源：[open enernoc data](https://open-enernoc-data.s3.amazonaws.com/anon/index.html)\
簡述：該資料集由EnerNOC電力公司提供，它依循時間序列記錄了2012年100棟不記名的商業大樓每5分鐘的用電情況。\
解釋變數：\
 - timestamp:  以秒為單位的絕對時間數值 (Type: int)\
 - dttm_utc: 即datetime，日期與時間 (Type: chr)\
 - value:  特定時間點下的電耗值 (Type: num)\
 - estimated: 若為1表示當下的電耗值為估計值，否則為0 (Type: chr)\ 
 - anomaly: 若有值表示當下的電耗值有誤，否則沒有值 (Type: chr)
 
## 2. 要用迴歸分析回答哪些問題
1. 電耗的日週期變化
2. 電耗的週週期變化
3. 預測兩週以內的電耗值
4. 預測三個月內的電耗
5. 預測半年的電耗

## 3. 使用的方法：
Generalized additive model (GAM)

## 4. 仔細的 model 解讀

```{r, message=FALSE, warning=FALSE}
library(feather)
library(data.table)
library(mgcv)
library(car)
library(ggplot2)
library(grid)
library(animation)
```

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = "D:\\WORKSPACE\\RProjects\\EnorNOC-GAM\\")
```

讀取資料
```{r}
library(data.table)
library(feather)
DT <- as.data.table(read_feather("D:/WORKSPACE/RProjects/EnorNOC-GAM/DT_4_ind"))
```

使用package `car`中的function `record`，將週次改為interger
```{r, echo=TRUE, message=FALSE, warning=FALSE, cache=TRUE, cache.vars='DT', results='hide'}
DT[, week_num := as.integer(car::recode(week,
    "'Monday'='1';'Tuesday'='2';'Wednesday'='3';'Thursday'='4';
    'Friday'='5';'Saturday'='6';'Sunday'='7'"))]
```

從讀取的資料中獲取industry, date, weekday and period等信息，並使用變量來儲存。
```{r}
n_type <- unique(DT[, type])
n_date <- unique(DT[, date])
n_weekdays <- unique(DT[, week])
period <- 48
```

截取兩個禮拜內的商業用樓房的電耗記錄，並儲存在data_r變量中。之後畫圖展示之。
```{r, fig.width = 8, fig.height=3.2}
data_r <- DT[(type == n_type[1] & date %in% n_date[57:70])]

ggplot(data_r, aes(date_time, value)) +
  geom_line() +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid.minor = element_line(colour = "grey90"),
        panel.grid.major = element_line(colour = "grey90"),
        panel.grid.major.x = element_line(colour = "grey90"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold")) +
  labs(x = "Date", y = "Load (kW)")
```

There is possible to see two main seasonalities in plotted time series: daily and weekly. We have 48 measurements during the day and 7 days during the week so that will be our independent variables to model response variable - electricity load. Let's construct it:
根據每天的週期性變化和每週的週期性變化，重新構建資料
```{r}
N <- nrow(data_r) # train set中的資料筆數
window <- N / period # train set所囊括的天數
matrix_gam <- data.table(Load = data_r[, value],
                         Daily = rep(1:period, window),
                         Weekly = data_r[, week_num])
```

Here we are! Train our first **GAM** with function `gam`. Independent variables are modeled by smoothing function `s`, for daily seasonality cubic regression spline is used, for weekly seasonality, P-splines is used, a number of knots are logically set to the number of unique values. Let's do it.
建立GAM模型，其中每天的週期性變化採用cubic regression spline模式來描述，每週的週期性變化採用P-splines來描述。
```{r}
gam_1 <- gam(Load ~ s(Daily, bs = "cr", k = period) +
               s(Weekly, bs = "ps", k = 7),
             data = matrix_gam,
             family = gaussian)
```

Package `mgcv` have many advantages and nice features. First is its visualization capabilities. Let's try it:
作圖分析此模型
```{r, fig.width = 8.5, fig.height=4}
layout(matrix(1:2, nrow = 1))
plot(gam_1, shade = TRUE)
```
在左圖中可以看出，用電高峰出現在下午3點；在右邊的圖中可以看出，週末的用電量小於平日。

查看模型的summary table
```{r}
summary(gam_1)
```
由summary table可以看出，根據p-value，電耗波動有很強的每天週期性變化和每週週期性變化。，Deviance explained數值不算很高，說明還有其他變量需要挖掘。

## 5. 預測
```{r}
data_test <- DT[(type == n_type[1] & date %in% n_date[71:84])]
matrix_gam <- data.table(Load = data_r[, value],
                         Daily = rep(1:period, window),
                         Weekly = data_test[, week_num])
pred_1 <- predict(gam_1, matrix_gam)
summary(pred_1)
plot(pred_1)
```

## 6. 課堂沒教的花俏的 r 指令
1. unique:獲取某欄位下的所有出現過的值
2. feather::read_feather: 讀取feather格式的資料
3. car::recode: 將資料重新編碼
